---
title: "Oct. 13th"
author: "Amy Atwater"
date: "October 13, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(curl)
curl("https://raw.githubusercontent.com/difiore/ADA2016/master/iqs.csv")
d <- read.csv(f, header=TRUE, sep=",", stringsAsFactors=FALSE)
head(d)
x <- d$IQ.before-d$IQ.after
m <- mean(x)
mu <- 0
s <- sd(x)
n <- length(x)
sem <- s/sqrt(n)
par(mfrow=c(1, 2))
boxplot(d$IQ.before, ylim=c(115,145), main="IQ",xlab="Before")
boxplot(d$IQ.after,ylim=c(115,145), main="IQ",xlab="After")
t <- (m - mu)/sem
t

alpha <- 0.05
crit <- qt(1 - alpha/2, df=n-1)
crit
test <- t < -crit || t > crit
test
t.test(x,df=n-1,alternative="two.sided")

pop <- c(rep(0,500),rep(1,500))
pi <- 0.5
x <- NULL
n <- 10
for (i in 1:1000){x[i] <- mean(sample(pop,size=n,replace=FALSE))}
m <- mean(x)
m
s <- sd(x)
s
pop_se <- sqrt(pi*(1-pi)/n) 
pop_se
v <- c(0,1,1,1,0,1,0,0,1,1,0,0,1,1,0,1,1,0,0,1,1,1,0,1,1,0,1,0,1,1)

### Challenge ###
### Null hypothesis is pi = 0.8
### alternative hypothesis is pi is less than 0.8
dicks <- mean(v)
dicks
pi <- 0.8
n <- 30
z <- (dicks - pi)/sqrt(pi*(1-pi)/30)
z
p <- pnorm(z, lower.tail=TRUE)
p
phat <- mean(v)
phat
pi <- 0.80
n <- 30
z <- (phat - pi)/sqrt(pi*(1-pi)/30)
z
p <- pnorm(z, lower.tail=TRUE)
p
lower <- phat + qnorm(0)*sqrt(phat*(1-phat)/30)
upper <- phat + qnorm(0.95)*sqrt(phat*(1-phat)/30)
ci <- c(lower, upper)
ci
pt <- prop.test(x=sum(v),n=length(v),p=0.8,conf.level=0.95,correct=FALSE, alternative = "less")
pt
### Example of the one sample test using proportion data (above) ###
typeI <- function(mu0,sigma,n,type="two-tailed",alpha=0.05,k=1000){p <- rep(NA,k) # sets up a vector of empty p values for(i in 1:k){ # sets up a loop to run k simulations
        x <- rnorm(n=n,mean=mu0,sd=sigma) # draws a sample from our distribution
        m <- mean(x) # calculates the mean
        s <- sd(x) # calculates the standard deviation
        t <- (m-mu0)/(s/sqrt(n)) # calculates the T statistic for the sample drawn from the null distribution relative to the null distribution
        if (type == "less") {
            p[i] <- pt(t,df=n-1,lower.tail=TRUE) # calculates the associated p value
        }
        if (type == "greater") {
            p[i] <- pt(t,df=n-1,lower.tail=FALSE)
        }
        if (type == "two-tailed") {
            if (t > 0) {p[i] <- 2 * pt(t,df=n-1,lower.tail=FALSE)}
            if (t < 0) {p[i] <- 2 * pt(t,df=n-1,lower.tail=TRUE)}
        }
    }
    curve(dnorm(x,mu0,sigma), mu0 - 4*sigma, mu0 + 4*sigma, main=paste("Type I = ",length(p[p<alpha])/k,sep=""), xlab="x", ylab="Pr(x)")
    if (type == "less") {
     polygon(cbind(c(mu0-4*sigma, seq(from=mu0-4*sigma, to=mu0-qnorm(1-alpha)*sigma, length.out=100), mu0-qnorm(1-alpha)*sigma)), c(0,dnorm(seq(from=mu0-4*sigma, to=mu0-qnorm(1-alpha)*sigma, length.out=100),mean=mu0,sd=sigma),0), border="black",col="gray")
    }
    if (type == "greater") {
     polygon(cbind(c(mu0+qnorm(1-alpha)*sigma, seq(from=mu0+qnorm(1-alpha)*sigma, to=mu0+4*sigma, length.out=100), mu0+4*sigma)), c(0,dnorm(seq(from=mu0+qnorm(1-alpha)*sigma, to=mu0+4*sigma, length.out=100),mean=mu0,sd=sigma),0), border="black",col="gray")
    }
    if (type == "two-tailed") {
     polygon(cbind(c(mu0-4*sigma, seq(from=mu0-4*sigma, to=mu0-qnorm(1-alpha/2)*sigma, length.out=100), mu0-qnorm(1-alpha/2)*sigma)), c(0,dnorm(seq(from=mu0-4*sigma, to=mu0-qnorm(1-alpha/2)*sigma, length.out=100),mean=mu0,sd=sigma),0), border="black",col="gray")
polygon(cbind(c(mu0+qnorm(1-alpha/2)*sigma, seq(from=mu0+qnorm(1-alpha/2)*sigma, to=mu0+4*sigma, length.out=100), mu0+4*sigma)), c(0,dnorm(seq(from=mu0+qnorm(1-alpha/2)*sigma, to=mu0+4*sigma, length.out=100),mean=mu0,sd=sigma),0), border="black",col="gray")} return(length(p[p<alpha])/k) # returns the proportion of simulations where p < alpha}

e <- typeI(mu0=-3,sigma=2,n=500,type="greater",alpha=0.05)
e
e
e <- typeI(mu0=-3,sigma=2,n=5,type="greater",alpha=0.05)
e
e <- typeI(mu0=-3,sigma=2,n=500,type="greater",alpha=0.5)
e
typeII <- function(mu0,muA,sigma,n,type="two-tailed",alpha=0.05,k=1000){
    p <- rep(NA,k) # sets up a vector of empty p values
    for(i in 1:k){
        x <- rnorm(n=n,mean=muA,sd=sigma) # draw from Ha
        m <- mean(x)
        s <- sd(x)
        t <- (m-mu0)/(s/sqrt(n)) # calculates the T statistic for the sample drawn from Ha relative to the null distribution
        if (type == "less") {
            p[i] <- pt(t,df=n-1,lower.tail=TRUE) # calculates the associated p value
        }
        if (type == "greater") {
            p[i] <- pt(t,df=n-1,lower.tail=FALSE)
        }
    if (type == "two.tailed") {
            if (t > 0) {p[i] <- 2 * pt(t,df=n-1,lower.tail=FALSE)}
            if (t < 0) {p[i] <- 2 * pt(t,df=n-1,lower.tail=TRUE)}
        }
    }

    curve(dnorm(x,mu0,sigma),mu0-4*sigma, mu0+4*sigma,
          main=paste("Type II - Prob of Incorrectly Failing To Reject Null\nBeta = ",length(p[p>=alpha])/k,sep=""), 
          xlab="x", ylab="Pr(x)",col="blue",lty=3,
          xlim=c(min(c(mu0-4*sigma, muA-4*sigma)),max(c(mu0+4*sigma,muA+4*sigma))),
          ylim=c(0,max(c(dnorm(mu0,mu0, sigma)),dnorm(muA,muA,sigma),dnorm(muA,muA,sigma/sqrt(n))))
          )
    curve(dnorm(x,muA,sigma),muA-4*sigma, muA+4*sigma, col="red", lty=3,add=TRUE)

    curve(dnorm(x,mu0,sigma/sqrt(n)),mu0-4*sigma/sqrt(n), mu0+4*sigma/sqrt(n),col="blue", add=TRUE)
    curve(dnorm(x,muA,sigma/sqrt(n)),muA-4*sigma/sqrt(n), muA+4*sigma/sqrt(n), col="red", add=TRUE)
    abline(h=0)

    if (type == "less") {
     polygon(cbind(
       c(mu0-qnorm(1-alpha)*sigma/sqrt(n), seq(from=mu0-qnorm(1-alpha)*sigma/sqrt(n),to=muA+4*sigma/sqrt(n),length.out=100), muA+4*sigma/sqrt(n))),
       c(0,dnorm(seq(mu0-qnorm(1-alpha)*sigma/sqrt(n), to=muA+4*sigma/sqrt(n), length.out=100), mean=muA,sd=sigma/sqrt(n)),0), border="red",col="red")
      abline(v=mu0-qnorm(1-alpha)*sigma/sqrt(n), col="black", lty=3, lwd=2)
    }

    if (type == "greater") {
     polygon(cbind(
       c(muA-4*sigma/sqrt(n), seq(from=muA-4*sigma/sqrt(n), to=mu0+qnorm(1-alpha)*sigma/sqrt(n), length.out=100), mu0+qnorm(1-alpha)*sigma/sqrt(n))),
       c(0,dnorm(seq(from=muA-4*sigma/sqrt(n), to=mu0+qnorm(1-alpha)*sigma/sqrt(n), length.out=100), mean=muA,sd=sigma/sqrt(n)),0), border="red",col="red")
      abline(v=mu0+qnorm(1-alpha)*sigma/sqrt(n), col="black", lty=3, lwd=2)
    } return(length(p[p>=alpha])/k)}

typeII <- function(mu0,muA,sigma,n,type="two-tailed",alpha=0.05,k=1000){
    p <- rep(NA,k) # sets up a vector of empty p values
    for(i in 1:k){
        x <- rnorm(n=n,mean=muA,sd=sigma) # draw from Ha
        m <- mean(x)
        s <- sd(x)
        t <- (m-mu0)/(s/sqrt(n)) # calculates the T statistic for the sample drawn from Ha relative to the null distribution
        if (type == "less") {
            p[i] <- pt(t,df=n-1,lower.tail=TRUE) # calculates the associated p value
        }
        if (type == "greater") {
            p[i] <- pt(t,df=n-1,lower.tail=FALSE)
        }
    if (type == "two.tailed") {
            if (t > 0) {p[i] <- 2 * pt(t,df=n-1,lower.tail=FALSE)}
            if (t < 0) {p[i] <- 2 * pt(t,df=n-1,lower.tail=TRUE)}
        }
    }

    curve(dnorm(x,mu0,sigma),mu0-4*sigma, mu0+4*sigma,
          main=paste("Type II - Prob of Incorrectly Failing To Reject Null\nBeta = ",length(p[p>=alpha])/k,sep=""), 
          xlab="x", ylab="Pr(x)",col="blue",lty=3,
          xlim=c(min(c(mu0-4*sigma, muA-4*sigma)),max(c(mu0+4*sigma,muA+4*sigma))),
          ylim=c(0,max(c(dnorm(mu0,mu0, sigma)),dnorm(muA,muA,sigma),dnorm(muA,muA,sigma/sqrt(n))))
          )
    curve(dnorm(x,muA,sigma),muA-4*sigma, muA+4*sigma, col="red", lty=3,add=TRUE)

    curve(dnorm(x,mu0,sigma/sqrt(n)),mu0-4*sigma/sqrt(n), mu0+4*sigma/sqrt(n),col="blue", add=TRUE)
    curve(dnorm(x,muA,sigma/sqrt(n)),muA-4*sigma/sqrt(n), muA+4*sigma/sqrt(n), col="red", add=TRUE)
    abline(h=0)

    if (type == "less") {
     polygon(cbind(
       c(mu0-qnorm(1-alpha)*sigma/sqrt(n), seq(from=mu0-qnorm(1-alpha)*sigma/sqrt(n),to=muA+4*sigma/sqrt(n),length.out=100), muA+4*sigma/sqrt(n))),
       c(0,dnorm(seq(mu0-qnorm(1-alpha)*sigma/sqrt(n), to=muA+4*sigma/sqrt(n), length.out=100), mean=muA,sd=sigma/sqrt(n)),0), border="red",col="red")
      abline(v=mu0-qnorm(1-alpha)*sigma/sqrt(n), col="black", lty=3, lwd=2)
    }

    if (type == "greater") {
     polygon(cbind(
       c(muA-4*sigma/sqrt(n), seq(from=muA-4*sigma/sqrt(n), to=mu0+qnorm(1-alpha)*sigma/sqrt(n), length.out=100), mu0+qnorm(1-alpha)*sigma/sqrt(n))),
       c(0,dnorm(seq(from=muA-4*sigma/sqrt(n), to=mu0+qnorm(1-alpha)*sigma/sqrt(n), length.out=100), mean=muA,sd=sigma/sqrt(n)),0), border="red",col="red")
      abline(v=mu0+qnorm(1-alpha)*sigma/sqrt(n), col="black", lty=3, lwd=2)
    }

return(length(p[p>=alpha])/k)
}


e <- typeII(mu0=2,muA=5,sigma=3,n=6,type="greater") # Ha > H0
e
e <- typeII(mu0=2,muA=1,sigma=3,n=6,type="less") # Ha < H0
e
e <- typeII(mu0=2,muA=1,sigma=5,n=6,type="less") # Ha < H0
e <- typeII(mu0=2,muA=5,sigma=3,n=6,type="less") # Ha < H0

ggplot(vis)
n <- 1:100
s <- seq(from = 1, to = 10, by = 1)
p <- matrix(data = NA, nrow = length(n), ncol = length(s))
p <- as.data.frame(p)
for (i in 1:length(n)) {for (j in 1:length(s)) {p[i, j] <- typeII(mu0 = 2, muA = 4, sigma = j, type = "two-tailed", alpha = 0.05, n = i)}}
p <- cbind(p, n)
p <- cbind(p,n)
manipulate(
  ggplot(data = p) + xlab("sample size n") + ylab("Type II Error Rate (Blue)\nand\nPower (Red)") + ylim(0,1) +
    geom_point(aes(x=n, y=p[,sigma]),colour="blue", alpha=1/2) + geom_line(aes(x=n, y=p[,sigma]),colour="blue", alpha=1/2) + geom_line(aes(x=n,y=0.8), colour="red", lty=3) + geom_point(aes(x=n, y=1-p[,sigma]), colour="red", alpha=1/2) + geom_line(aes(x=n, y=1-p[,sigma]),colour="red", alpha=1/2), sigma=slider(1, 10, initial=3, step=1)
)

### MODULE 12 ###
library(curl)
library(ggplot2)
library(gridExtra)
library(manipulate)
f <- f <- curl("https://raw.githubusercontent.com/difiore/ADA2016/master/zombies.csv")
d <- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(d)
plot(data = d, height ~ weight)
